### 2013-12-14 paper/paper.tex
Add paper in TeX.

% THIS IS SIGPROC-SP.TEX - VERSION 3.1
% WORKS WITH V3.2SP OF ACM_PROC_ARTICLE-SP.CLS
% APRIL 2009
%
% It is an example file showing how to use the 'acm_proc_article-sp.cls' V3.2SP
% LaTeX2e document class file for Conference Proceedings submissions.
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V3.2SP) *DOES NOT* produce:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) Page numbering
% ---------------------------------------------------------------------------------------------------------------
% It is an example which *does* use the .bib file (from which the .bbl file
% is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission,
% you need to 'insert'  your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% Questions regarding SIGS should be sent to
% Adrienne Griscti ---> griscti@acm.org
%
% Questions/suggestions regarding the guidelines, .tex and .cls files, etc. to
% Gerald Murray ---> murray@hq.acm.org
%
% For tracking purposes - this is V3.1SP - APRIL 2009

\documentclass{acm_proc_article-sp}

\begin{document}

\title{Market-based Mechanisms for Resource Scheduling}
%\subtitle{[Extended Abstract]
%\titlenote{A full version of this paper is available as
%\textit{Author's Guide to Preparing ACM SIG Proceedings Using
%\LaTeX$2_\epsilon$\ and BibTeX} at
%\texttt{www.acm.org/eaddress.htm}}}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

\numberofauthors{3} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Paul Christiano
% 2nd. author
\alignauthor
Frank Li
% 3rd. author
\alignauthor
Richard Shin
%\and  % use '\and' if you need 'another row' of author names
%% 4th. author
%\alignauthor Lawrence P. Leipuner\\
%       \affaddr{Brookhaven Laboratories}\\
%       \affaddr{Brookhaven National Lab}\\
%       \affaddr{P.O. Box 5000}\\
%       \email{lleipuner@researchlabs.org}
%% 5th. author
%\alignauthor Sean Fogarty\\
%       \affaddr{NASA Ames Research Center}\\
%       \affaddr{Moffett Field}\\
%       \affaddr{California 94035}\\
%       \email{fogartys@amesres.org}
%% 6th. author
%\alignauthor Charles Palmer\\
%       \affaddr{Palmer Research Laboratories}\\
%       \affaddr{8600 Datapoint Drive}\\
%       \affaddr{San Antonio, Texas 78229}\\
%       \email{cpalmer@prl.com}
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
%\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
%email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
%(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
%\date{30 July 1999}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

\maketitle
\begin{abstract}
Large-scale computing clusters offer potential for greater efficiency through
pooling of computational resources, but must contend with the problem of
allocating these resources to users with competitive and potentially
incompatible demands. Such conflicts are typically arbitrated \emph{fairness}
criterions, such as dominant resource fairness (DRF), which benefit from their
simplicity, efficiency, and non-manipulability. Unfortunately, when applied in
realistic settings---when dealing with inhomogeneous resources, time-varying
demands, requirements for data locality, etc.---these systems cease to be either
efficient or non-manipulable.

We propose using prices as an abstraction for subsystems to communicate their
constraints, and describe a protocol based on auctions for allocating resources
in a cluster. Our protocol has the key advantage that it can continue to meet
efficiency guarantees in realistic environments where preferences are
inhomogeneous across resources and time, and in the presence of adversarial
manipulation. Moreover, we propose prices as a general mechanism for enforcing a
clean abstraction between different components of a system, allowing one
component to make choices intelligently based on constraints elsewhere in the
system even when it was not designed with knowledge of those constraints.

We implement these ideas in the Mesos cluster manager, along with microeconomic
``agents'' which allow traditional Mesos clients to interact seamlessly with our
modified implementation. We show that the computational and communication
overhead proposed by our system is manageable and that its performance compares
favorably to existing approaches (particularly DRF) even when users don’t
provide any meaningful price information. Moreover, we show that when accurate
economic data is available our system is able to achieve substantially higher
social welfare than DRF. Finally, we discuss a number of examples where this
functionality may be useful even in applications where the end-users of the
system don't have to think about economics at all, and instead prices are
generated algorithmically as a principled way to accommodate constraints that
are currently managed in an ad hoc way.
\end{abstract}

% A category with the (minimum) three required fields
%\category{H.4}{Information Systems Applications}{Miscellaneous}
%A category including the fourth, optional field follows...
%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

%\terms{Theory}

%\keywords{ACM proceedings, \LaTeX, text tagging} % NOT required for Proceedings

\section{Introduction}
\subsection{Overview}
We consider an environment in which a single cluster is used for a variety of
tasks, which may have very different demands for resources. For example, a
company might run research tasks in the same cluster as production web services,
in order to make use of computational resources which would otherwise lie idle
when web traffic is low. In general, running different tasks in the same cluster
has the promise of improving utilization and reducing overhead.

Different tasks may have significantly different demands for resources; some may
have deadlines on the scale of seconds while others have deadlines of hours or
days, some may require specialized hardware available on only certain machines,
some may require using the same machines consistently while others can use
different machines at different times, some may have complex constraints on data
locality, etc.

\subsection{Mesos}
The Mesos cluster manager is designed to address these problems, and to run
several independent \emph{frameworks} on a single cluster. Its core design philosophy
is to make minimal assumptions about the behavior of the frameworks it hosts. To
this end it allocates resources using ``resource offers,'' and leaves it up to the
frameworks themselves to decide which resources they will use and which they
will pass on. Frameworks are free to make these decisions on the basis of the
detailed specifications of the tasks they want done (for example, to accept
machines that improve data locality), but Mesos itself makes use of a very
simple fairness criterion to decide which frameworks receive offers.

The abstraction of resource offers allows Mesos to serve frameworks without
understanding the constraints they face. By making offers to those frameworks
which are currently using the least resources, Mesos attempts to ensure that all
frameworks receive a fair share of the cluster. Because of the simplicity of
this protocol, it appears to be immune to manipulation (and indeed it has been
shown to be immune to manipulation in theory).

Unfortunately, while the fairness policy used by Mesos is efficient and
strategy-proof in a simple one-shot game with homogeneous resources, it is
neither efficient nor strategy-proof in this setting. For example, a framework
using Mesos which prefers to run tasks on machines that already have relevant
data is forced to engage in strategic behavior: given a resource which it
\emph{could} use now, should it take the resource or decline it, so that it
might receive a better offer in the future? Solutions to this problem rely on
simple heuristics which require tuning based on the parameters of the system;
there are simple examples where the resulting allocations are not efficient
despite best-efforts of the clients, and in complex examples there is little to
believe the results are efficient. Moreover, despite being ``strategy-proof'' the
current design of Mesos is essentially unsuitable for even mildly hostile
environments; a framework which can use less than its ``fair share'' of the
resources at any point in time has no incentive to do so, and indeed has every
incentive to exploit all resources it can be allocated even if they are being
used arbitrarily inefficiently (and there is always \emph{something} to do with
idle resources, if only perform proof-of-work in exchange for small amounts of
money).

\subsection{Microeconomic methods}
In this paper we confront the challenge of modifying Mesos to provide more
robust efficiency guarantees, without compromising its agnosticism regarding the
nature of different frameworks’ desiderata. We propose \emph{price signals} as a
mechanism for accomplishing this; instead of accepting or rejecting a resource
offer, a framework can express its preferences by assigning a \emph{willingness
to pay} for that resource, and Mesos can give the resources to whichever
frameworks are most willing to pay.

We stress the distinction between this use of price signals and the more typical
motivation. We are most interested in price signals as an \emph{abstraction} which
programs can use to express their constraints, rather than as a mechanism to
extract revenue for the cluster operator or to arbitrate between competing
external interests. To consider a simple case, suppose there are two programs
that are contending for a shared resource $R$. Each could either make use of $R$ or
some alternative resource $R'$ (for example, a different machine in the cluster at
a different time). Our proposal is to have each program articulate the value of
resource $R$ (compared to the benchmark of not receiving resource $R$, and then
contending for resource $R'$ at a later date). The primary alternatives that have
been considered are:
\begin{enumerate}
  \item Ignore the relative value of resource $R$ and $R'$, and make allocation
    decisions entirely on the basis of fairness considerations (for example by
    ensuring that each framework has an equal number of every resource). This is
    the approach taken by Mesos, although it in fact achieves better behavior
    precisely by \emph{violating} the strategy-proofness of its underlying allocation
    policy, and providing incentives to frameworks to strategically decline a
    resource in order to receive a better allocation in the future.
  \item Understand the entire optimization problem faced by each problem, and
    make explicit judgments of the relative value of $R$ and $R'$ within the
    scheduler. This violates the spirit of Mesos, and in general appears to be
    difficult given the complexity and scale of the cluster scheduling problem.
\end{enumerate}

\subsection{Guide to the paper}
In section 2, we make the theoretical case for using microeconomic methods. We
provide a litany of examples in which existing approaches, and in particular
DRF, fail to be either efficient or strategy-proof, and describe examples of
strategic behavior in Mesos. We discuss a number of examples in which economic
data generated \emph{within} the system could be useful, and compare this approach to a
number of ad hoc approaches currently pursued to solve similar problems.

In section 3, we describe our protocol and the most important design decisions
we made. In particular, we describe the mechanisms we used to ensure that our
protocol remains backwards-compatible, and can be applied without any change to
the clients’ code.

In section 4, we describe our implementation of this protocol within Mesos.

In section 5, we report on experimental comparisons between our results and
existing approaches: (a) we evaluate the computational and communication
overhead of our approach and show that it is manageable, (b) we compare our
system in ``backwards-compatibility mode,'' without any economic information
whatsoever, to DRF, and show that it performs comparably whether judged by
utilization or fairness, (c) we examine a number of cases in which economic
information is available; we show that in general our algorithms achieve higher
social efficiency than DRF, and provide natural examples in which DRF fails to
be Pareto efficient.

\section{The case for microeconomic scheduling}
\subsection{Scheduling is hard}
Consider the following typical situations within a cluster:
\begin{enumerate}
  \item Some clients have regular throughput and flexible deadlines, while some
    clients have hard real-time constraints and variable traffic.
  \item Different tasks make use of different characteristics of hardware on
    machines in the cluster, for example high ratios of RAM to CPU on a
    particular machine, or GPUs or other hardware that is available on only
    certain machines.
  \item Tasks require data which is located on disk or even in memory somewhere
    in the cluster, and co-locating the tasks with the data improves
    performance.
\end{enumerate}
In each of these examples, clients of the cluster have preferences over
allocations which are difficult to precisely express. It would be possible but
difficult to cover just these cases, but the reader can doubtless generate many
similar cases, and as the environment changes the nature of these constraints is
likely to continue to change as well.

Mesos is able to schedule even in the face of such constraints, by making
resource offers to clients and allowing them to accept or reject resources
according to whether they satisfy their constraints. But this is approach offers
little granularity to clients, forces them to adopt brittle heuristics, and
moreover compromises the guarantees of the fairness policies Mesos uses to
decide how to make resource offers.

Mesos uses a fair-share scheduler based on dominant resource fairness. This
means that the client to receive the next resource offer is the one who is
currently using the minimum share of the resource they are using the \emph{most}
of (relative to the total supply in the cluster). A weighted version can be
defined analogously.

Consider a client of Mesos who is interested in receiving a machine with
characteristic X, because it would allow their task to run 20\% faster (or
perhaps because it would allow them to run a task which is slightly higher
priority than the task they would run on a machine without characteristic X).
Given an offer for a machine without characteristic X, what should this client
do? The formal guarantee of ``strategy-proofness'' suggests the client would take
it; indeed, to do otherwise would explicitly be strategic. But in fact, we hope
that clients will yield a highly-contended resource when they can use an
alternative, or will yield a typical resource if they could make particularly
good use of an alternative. And indeed this is done by many clients to Mesos.

This example clearly shows that in this case, the mechanism provided by Mesos is
\emph{not} strategy-proof. Worse, it’s not Pareto efficient if people are
honest, and need not be Pareto efficient if clients behave optimally. Similar
things happen in the other examples we mentioned, or in most of the similar
alternatives which we can envision:
\begin{enumerate}
  \item When one client has a constant stream of work and another has varying
    traffic, Mesos encourages both to constantly use up their allocated share of
    the network (unless they have nothing to do with the space). But an
    efficient allocation would give the client with varying workload more
    resources when it has more work. 
  \item Mesos gives a valuable resource to whatever client happens to be using
    the fewest resources when that machine becomes available. This encourages
    clients to inefficiently decline resources which they could use, so that
    they will be offered better resources in the future. This may decrease
    utilization, and moreover it provides little guarantee that a valuable
    resource will be used by the framework which most values it---instead it will
    be used by whichever framework uses the least resources.
  \item If a client is waiting for a machine that is co-located with some data
    that their task needs, then Mesos encourages them to guess how likely that
    resource is to be freed up soon, and to decline new resources if this
    probability is high enough. This is actually much more efficient behavior
    than if the client behaved honestly by accepting whatever resources it could
    use, but it provides some indication that the ``guarantees'' of DRF are not
    directly applicable to the situations Mesos is intended to cover.
  \end{enumerate}

These shortcomings are not specific to DRF; these will be issues for any
scheduler which treats the scheduling problem as a one-shot game with
homogeneous resources when in fact it is more complicated. Theoretical
guarantees are not likely to be meaningful in this regime. To evaluate proposed
solutions, we suggest a combination of:
\begin{enumerate}
  \item Theoretical analyses which apply in realistic contexts
  \item Experimental evidence for effectiveness
\end{enumerate}
We provide strong evidence that our proposal is at least not inferior to DRF on
either of these grounds, and we provide some evidence that even our highly
preliminary implementation is already superior.

\subsection{Existing approaches}
Many difficulties with scheduling have been isolated and addressed, typically by
ad hoc mechanisms which require tuning and don't necessarily play nicely with
each other. In order to contrast these approaches with microeconomic methods, we
list some examples of solutions to these problems here. Note that we will not be
able to solve all of these problems within our proposal---that would be an
extraordinarily ambitious project. We list these examples to give some sense of
how a microeconomic framework could be used to replace ad hoc approaches,
improving robustness and ease-of-design. In contrast, our goal in this paper is
simply to show that microeconomic approaches really are a feasible option, and
that in the ``base case'' they are as good as existing solutions.

\paragraph{Deadlines} DRF makes no provision for coping with deadlines, which
are often of considerable practical importance. The Jockey system enforces
deadlines in the context of a fair-share scheduler, by increasing or decreasing
a task's share of the cluster depending on the importance of their deadline.
While this is a significant improvement over the alternative, we find it a very
crude solution to the underlying problem:
\begin{enumerate}
  \item Jockey has no dependence on how important resources currently are for
    the cluster, it just assigns a job as many resources as are needed to meet
    its deadline.
  \item It is not clear how to generalize Jockey to systems where multiple
    simultaneous tasks have deadlines, or where meeting deadlines must be
    balanced against any other consideration.
  \item Jockey changes the allocations of tasks depending on their performance,
    and will typically invalidate any guarantees provided by the original
    scheduler.
  \item Jockey does not have any principled approach for reasoning about the
    error bars on a task's completion time (instead this is determined by a
    fine-tuned parameter of the system).
  \item Jockey cannot reason about multiple deadlines of varying importance, but
    instead gives a job as many resources as it needs to finish as soon as it
    would be useful.
\end{enumerate}

The microeconomic alternative makes use of Jockey's core component, an estimate
for the time required to complete a task, but uses a different control system.
In the microeconomic framework a deadline is associated with an economic value
(that might be algorithmically generated based on the priority of the task)
which can be compared to the economic value of other activity in the cluster.
The importance of resources to the task is then given by the effect those
resources will have on the task's probability of completing by deadline, and
this can be compared to the importance of those resources to other processes
running on the cluster. Resources can then be shifted until these two values are
equal.

\paragraph{Priority and preemption} In existing systems priority is typically
implemented as an absolute ordering; a task of high priority can preempt a task
of lower priority and always wins in contention for resources. This approach is
conceptually problematic because there are always mechanisms for a
higher-priority process to improve performance by increasing its resource usage
(for example speculative execution, or increased duplication of straggling
processes) and an infinite tradeoff between the performance of high-priority and
low-priority processes is undesirable. In practice, this approach relies on
``reasonable'' behavior by high-priority processes, which makes reasoning about
systems difficult. Such systems are also brittle and prone to priority
inversions, in which low-priority processes can be inadvertently elevated to
high-priority.

Another approach is to assign tasks of different types appropriate shares of a
cluster. But this approach also fails to capture what we really care about. For
example, we would like high priority processes to expand their use of the
cluster when they need it.

These problems are considerably exacerbated when we consider preemption,
especially if preemption is costly for the preempted process: when should a low
priority process be allowed to finish, and when should it be preempted?

Microeconomic methods allow for a principled answer to these questions: by
describing ``high priority'' tasks by \emph{how much more} we care about them
finishing faster, we can make principled tradeoffs.

\paragraph{Data locality} Many tasks run more efficiently when they are
co-located with data. At the moment, this is typically achieved either by
explicit optimization (requiring the scheduler to be aware of detailed
characteristics of the tasks its scheduling) or by heuristics, for example
blocking for 5 seconds before accepting an offered resource which is not local
to data. Again, microeconomic methods provide a principled approach which is
compatible with a lightweight interface between client and scheduler: a client
can estimate how much more slowly a task will run on a machine without the data,
can estimate how likely a machine with the data is to become free, and can
thereby estimate how much of a discount they would need to be offered before
they would accept a machine without their data.

\subsection{Philosophical motivation}
Our primary motivation is strongly to maximize efficiency, rather than other
measures such as utilization, fairness, or strategy-proofness. We view other
characteristics as necessary insofar as they help achieve efficiency.

In some sense this is problematic, because efficiency is typically unmeasurable.
But there are a number of pragmatically relevant implications:
\begin{enumerate}
  \item Pareto efficiency is a minimal standard for acceptability. To the extent
    that existing approaches fail to be Pareto efficient, this is much more
    damning than a possible failure to be strategy-proof (as long as the
    possibility of manipulation does not in itself destroy our efficiency
    guarantee!)
  \item Even where true prices are unavailable, we think in terms of the real,
    unobserved value of an outcome (such as finishing a computation or meeting a
    deadline). Many existing schemes need to implicitly make judgments about
    these values (because they are making tradeoffs between different outcomes).
    So we don't treat the difficulty of assigning values to outcomes as a
    deal-breaker; to the extent that we already need to make implicit tradeoffs,
    reifying those tradeoffs is only likely to help us improve the situation.
\end{enumerate}

\subsection{Why microeconomic methods?}
The basic paradigm we propose is conceptually simple, though is associated with
many practical difficulties. At each stage where we need to make a decision, we
propose:
\begin{enumerate}
  \item Estimate the value of different outcomes.
  \item Estimate the outcomes that will result from different decisions.
  \item Compute the value of different decisions.
\end{enumerate}

This paradigm has the virtue of being modular in a number of important respects:
\begin{enumerate}
  \item The notion of value provides a unifying framework for different
    components to communicate with each other; if I am one component providing a
    service to some client, and I know how valuable different outcomes are to my
    client, I know everything I need to know; my policy does not need to depend
    on anything else about the client.
  \item I can improve and reason about the prediction of outcomes and the
    estimates of value separately. 
  \item It is often possible to compute the value straightforwardly, given the
    values stated  by my clients. For example, if I am scheduling a MapReduce
    instance and know that a 1 second delay costs my client 1 unit, then I can
    estimate the cost of a 1 second delay in one of my tasks simply by
    estimating (or using a crude heuristic for) the probability that this task
    is on the critical path.
\end{enumerate}

%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional
%\balancecolumns
%\appendix
%Appendix A
%\section{Headings in Appendices}
%The rules about hierarchical headings discussed above for
%the body of the article are different in the appendices.
%In the \textbf{appendix} environment, the command
%\textbf{section} is used to
%indicate the start of each Appendix, with alphabetic order
%designation (i.e. the first is A, the second B, etc.) and
%a title (if you include one).  So, if you need
%hierarchical structure
%\textit{within} an Appendix, start with \textbf{subsection} as the
%highest level. Here is an outline of the body of this
%document in Appendix-appropriate form:
%\subsection{Introduction}
%\subsection{The Body of the Paper}
%\subsubsection{Type Changes and  Special Characters}
%\subsubsection{Math Equations}
%\paragraph{Inline (In-text) Equations}
%\paragraph{Display Equations}
%\subsubsection{Citations}
%\subsubsection{Tables}
%\subsubsection{Figures}
%\subsubsection{Theorem-like Constructs}
%\subsubsection*{A Caveat for the \TeX\ Expert}
%\subsection{Conclusions}
%\subsection{Acknowledgments}
%\subsection{Additional Authors}
%This section is inserted by \LaTeX; you do not insert it.
%You just add the names and information in the
%\texttt{{\char'134}additionalauthors} command at the start
%of the document.
%\subsection{References}
%Generated by bibtex from your ~.bib file.  Run latex,
%then bibtex, then latex twice (to resolve references)
%to create the ~.bbl file.  Insert that ~.bbl file into
%the .tex source file and comment out
%the command \texttt{{\char'134}thebibliography}.
%% This next section command marks the start of
%% Appendix B, and does not continue the present hierarchy
%\section{More Help for the Hardy}
%The acm\_proc\_article-sp document class file itself is chock-full of succinct
%and helpful comments.  If you consider yourself a moderately
%experienced to expert user of \LaTeX, you may find reading
%it useful but please remember not to change it.
\balancecolumns
% That's all folks!
\end{document}